# *Ecological Archives* Exxxxx
# Fuhr M. *et al* Long-term stem inventory data from mountain forest plots in France. *Ecology* xxxx


```{r, results="hide", echo=FALSE, message=FALSE}
knitr::opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap="")
options(citation_format = "text", max.names=2)
require(bibtex)
bib <- read.bibtex("refs.bib")
source(file.path('..', 'R', 'ReadData.R'))
data_c <- readRDS('../output/data_c.rds')
data_m <- readRDS('../output/data_m.rds')
data_p <- readRDS('../output/data_p.rds')
df_growth <- readRDS('../output/df_growth.rds')
```

```{r precite, results='hide', message=FALSE, echo=FALSE}
# Kluge to avoid all author names being cited. Bug in knitcitations, see:
# https://github.com/cboettig/knitcitations/issues/65
```


## INTRODUCTION

TODO `r citep(list(bib["zianis_biomass_2005"]))`.


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
# TO IMPROVE
map_plots(data_p)
```
**Figure 1** Plots distribution in France.

 [github](https://github.com/dfalster/baad/).


## METHODS

## METADATA CLASS I. DATA SET DESCRIPTORS

### A. DATA SET IDENTITY
Long-term stem inventory data from mountain forest plots in France.

### B. DATA SET IDENTIFICATION CODE:

### C. DATA SET DESCRIPTORS:
**1. Originators**: The study was initiated by TODO.

**2. Abstract**: We present repeated stem measurement data from `r length(unique(data_p$site_id))` plots in mountain forests in France.

### D. Key words

France; mortality; permanent plot; mountain forest; recruitment; stem growth; stem height; canopy radius and height; forest management


## METADATA CLASS II. RESEARCH ORIGIN DESCRIPTORS

### A. "OVERALL" PROJECT DESCRIPTION

**1. Identity**: 

**2. Originators**: 

**3. Period of study**: From `r min(data_p$year_first_mes, na.rm = TRUE)` to 2015.

**4. Objectives**:

**5. Abstract**:

**6. Sources of funding**:
	

### B. SPECIFIC SUBPROJECT DESCRIPTION

#### II.B.1. Census of stems ≥7.5 cm DBH

**1. Sites description**: TODO site type, geography, Habitat, Geology Landform, Site history, Climate

**2. Experimental or sampling design**: TODO 

**3. Research methods**: TODO Field, Taxonomy, Permit history, Legal


## CLASS III. DATA SET STATUS AND ACCESSIBILITY

### A. Status

**Latest update**: The dataset made available through this publication was processed on TODO. Newer updates may be retrieved from the following website TODO [github.com/kunstler](https://github.com/kunstler).

**Metadata status**: Metadata are complete to last update and are stored with data.

**Data verification**: TODO

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
# TO IMPROVE
plot_growth_error(df_growth)
```
**Figure 2** Growth in function of iniotial dbh. and outlier identification.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=7}
# TO IMPROVE
plot_allo_error(data_m)
```
**Figure 3** Tree allometry and outlier identification.


### B. Accessibility

**1. Storage location and medium**: The data published on Ecological Archives is the first release stored at [github.com/kunstler](https://github.com/kunstler). 

If you notice any problems with the dataset, please list an issue at [github.com/kunstler](https://github.com/kunstler).

**Contact persons**: TODO.

**Copyright restrictions**: The dataset is released under the the [Creative Commons Zero](https://creativecommons.org/publicdomain/zero/1.0/) ??.

**Costs**: None.

## CLASS IV. DATA STRUCTURAL DESCRIPTORS

### A. Data Set Files

The following files are provided:

- `data_c.csv`: Tree coordinates file
- `data_m.csv`: Tree measurements file
- `data_p.csv`: Plots description file
- `metadata_data_c.csv`: Metatdata for tree coordinates file
- `metadata_data_m.csv`: Metatdata for tree measurements file
- `metadata_data_p.csv`: Metatdata for plots description file


### B. Variable definitions for data.csv

```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
metadata_data_c <- read.csv(file.path('..', 'output', 'metadata_data_c.csv'))
kable(metadata_data_c, "markdown", align="l", padding=10)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
metadata_data_m <- read.csv(file.path('..', 'output', 'metadata_data_m.csv'))
kable(metadata_data_m, "markdown", align="l", padding=10)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, comment='', results="asis"}
metadata_data_p <- read.csv(file.path('..', 'output', 'metadata_data_p.csv'))
kable(metadata_data_p, "markdown", align="l", padding=10)
```

**Number of observations by files**

```{r results='asis', echo=FALSE, warning=FALSE}
options(digits=9)
data_summary <- data.frame(file = c('data_c', 'data_m', 'data_p'),
                           Nobs = c(nrow(data_c), nrow(data_m), nrow(data_p))) 
kable(data_summary, "markdown", align="l", padding=10)
```

## ACKNOWLEDGMENTS

**Compilers:** TODO.

## REFERENCES

